A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-01-23 paper/paper.tex
paper: fix some underfull hboxes, rephrase "see eg."

  This paper describes an uniform indexed characterization of both augmented semi-simplicial and semi-cubical sets arising respectively as unary and binary iterated parametricity-based constructions.

@@

\section{Introduction}
In the context of functional programming, Reynolds' parametricity~\cite{reynolds72} interprets types as relations characterizing the observational behavior of programs of this type. Parametricity can be iterated, and it has been noted that iterated Reynolds' parametricity, when using more generally graphs than relations, has a cubical flavor (see e.g.~\cite{johann17,altenkirch15,moulin16,moeneclaey21,moeneclaey22phd}). We obtain a \emph{unary} variant of Reynolds' \emph{binary} parametricity by using predicates or families instead of relations or graphs, in which case, we obtain a form of realizability (see e.g.~\cite{bernardy12,lasson12,moulin16}). It has then been noted that iterated unary parametricity has an augmented simplicial flavor~\footnote{Private communication with Thorsten Altenkirch and Hugo Moeneclaey}. % TODO anonymize??

We exploit the connection between iterated unary parametricity and augmented simplicial sets, and between iterated binary parametricity and cubical sets to present a uniform construction of both augmented semi-simplicial and semi-cubical sets obtained by applying simple rules characterizing iterated parametricity. In contrast to the standard presheaf presentation of augmented semi-simplicial and semi-cubical sets (see e.g.~\cite{fri08,grandis03,buchholtz17}), our uniform construction, which we call \emph{$\nu$-sets}, is \emph{indexed}. That is, instead of having the set of augmented semi-simplices or semi-cubes in dimension $n+1$ \emph{fibered} over the set of augmented semi-simplices or semi-cubes in dimension $n$, we consider families of augmented semi-simplices and semi-cubes indexed over their faces.

@@

Since we have formalized the proof in the Coq proof assistant, which implements an intensional type theory, we now present the tables in \emph{intensional} type theory~\cite{martin-lof...,Coq}, leaving implicit some details, which we will explain here, and other details, which we will explain in \ref{sec:mech}.

>>>>

  This paper describes a uniform indexed characterization of both augmented semi-simplicial and semi-cubical sets arising respectively as unary and binary iterated parametricity-based constructions.

@@

\section{Introduction}
In the context of functional programming, Reynolds' parametricity~\cite{reynolds72} interprets types as relations characterizing the observational behavior of programs of this type. Parametricity can be iterated, and it has been noted that iterated Reynolds' parametricity, when using more generally graphs than relations, has a cubical flavor~\cite{johann17,altenkirch15,moulin16,moeneclaey21,moeneclaey22phd}. We obtain a \emph{unary} variant of Reynolds' \emph{binary} parametricity by using predicates or families instead of relations or graphs, in which case, we obtain a form of realizability~\cite{bernardy12,lasson12,moulin16}. It has then been noted that iterated unary parametricity has an augmented simplicial flavor~\footnote{Private communication with Hugo Moeneclaey and Thorsten Altenkirch}. % TODO anonymize??

We exploit the connection between iterated unary parametricity and augmented simplicial sets, and between iterated binary parametricity and cubical sets to present a uniform construction of both augmented semi-simplicial and semi-cubical sets obtained by applying simple rules characterizing iterated parametricity. In contrast to the standard presheaf presentation of augmented semi-simplicial and semi-cubical sets~\cite{fri08,grandis03,buchholtz17}, our uniform construction, which we call \emph{$\nu$-sets}, is \emph{indexed}. That is, instead of having the set of augmented semi-simplices or semi-cubes in dimension $n+1$ \emph{fibered} over the set of augmented semi-simplices or semi-cubes in dimension $n$, we consider families of augmented semi-simplices and semi-cubes indexed over their faces.

@@

Since we have formalized the proof in the Coq proof assistant, which implements an intensional type theory, we now present the tables in \emph{intensional} type theory~\cite{martinlof75,martinlof84,coq22}, leaving implicit some details, which we will explain here, and other details, which we will explain in \ref{sec:mech}.
